Random Walks across Social Networks: The Markovian Movement of a Stochastic Soccerball

### 2020-07-16 rmd/analysis_in_R.Rmd
writing draft

---
jupyter:
  jupytext:
    formats: ipynb,Rmd
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

# Executive Summary

# Introduction

```{r}
library(plyr) # data manipulation @HWickham2009
library(dplyr) # data manipulation @HWickham@2017
library(knitr) # dynamic report generation @YXie2012
library(igraph) # for graph/network analysis @GCsardiTNepusz2006
library(ggraph) # graph visualization in the style of ggplot @TPederson2016a
library(netrankr) # network centrality @DSchoch2015
library(ggplot2) # for plotting visualizations @HWickham2007
library(ggforce) # adds functionality to ggplot2, eg. geom_circle @TPederson2016b
library(ggrepel) # adds 'repellent' non-overlapping labels to ggplot @KSlowikowski2016
library(pheatmap) # prettier heat maps
library(RColorBrewer) # color palettes
library(docstring) # allows ?help to display function descriptions analogous to Pythonic docstrings
```

```{r}
DrawPitch <- function(lengthPitch=105, widthPitch=68, arrow=c("none", "r", "l"), theme=c("light", "dark", "grey", "grass")) {
    #' Draw regulation football pitch with penalty areas and centre circle.
    #' Adapted from https://github.com/JoGall/soccermatics

@@

  border <- c(5, 5, 5, 5)

@@

  # add arrow
  if(arrow[1] == "r") {
    p <- p + 
      geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  } else if(arrow[1] == "l") {
    p <- p + 
      geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), 
                   colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
  }

  return(p)
}
```

# Data

```{r}
LoadAndScaleData <- function(national_league) {
    #' Load preprocessed event logs for given national league.

    events <- read.csv(paste0('../data/processed/', national_league, '.csv'))

@@

    events$location_y <- y_scale - (events$location_y/100 * y_scale)
    events$target_y <- y_scale - (events$target_y/100 * y_scale)

    return(events)
}

@@

```{r}
events <- LoadAndScaleData('England')
```

```{r}
events[1:10,]
```

```{r}
LeagueResults <- function(league){
    #' Calculate results and table for given national league.

    teams <- sort(unique(league$team))

    home_score <- matrix(0, nrow=length(teams), ncol=length(teams))
    colnames(home_score) <- teams
    rownames(home_score) <- teams

    away_score <- home_score

    league_table <- matrix(0, nrow=length(teams), ncol=9)
    rownames(league_table) <- teams
    colnames(league_table) <- c('Pos','Pld','W','D','L','GF','GA','GD','Pts')

    for (m in unique(league$matchId)){

        match <- league[league$matchId==m,]
        home_team <- unique(match[match$home_or_away=='Home',]$team)
        away_team <- unique(match[match$home_or_away=='Away',]$team)
        home_goals <- nrow(match[match$to_team=='Home Goal',])
        away_goals <- nrow(match[match$to_team=='Away Goal',])

        home_score[home_team, away_team] <- home_goals
        away_score[home_team, away_team] <- away_goals

        if (home_goals > away_goals){
            league_table[home_team,'W'] <- league_table[home_team,'W'] + 1
            league_table[away_team,'L'] <- league_table[away_team,'L'] + 1
        } else if (away_goals > home_goals){
            league_table[home_team,'L'] <- league_table[home_team,'L'] + 1
            league_table[away_team,'W'] <- league_table[away_team,'W'] + 1    
        } else if (home_goals == away_goals) {
            league_table[home_team,'D'] <- league_table[home_team,'D'] + 1
            league_table[away_team,'D'] <- league_table[away_team,'D'] + 1
        }

        league_table[home_team,'GF'] <- league_table[home_team,'GF'] + home_goals
        league_table[away_team,'GF'] <- league_table[away_team,'GF'] + away_goals
        league_table[home_team,'GA'] <- league_table[home_team,'GA'] + away_goals
        league_table[away_team,'GA'] <- league_table[away_team,'GA'] + home_goals

        for (team in c(home_team, away_team)){
            league_table[team,'GD'] <- league_table[team,'GF'] - league_table[team,'GA']
            league_table[team,'Pts'] <- 3*league_table[team,'W'] + 1*league_table[team,'D']
            league_table[team,'Pld'] <- league_table[team,'Pld'] + 1
        }

    }

    league_table <- league_table[order(-league_table[,'Pts']),]

    league_table[,'Pos'] <- 1:20

    return(league_table)
}
```

```{r}
prem <- LeagueResults(events)
```

```{r}
pl <- as.data.frame(prem)
```

```{r}
pl['Team'] <- rownames(pl)
```

```{r}
pl2 <- as.data.frame(pl[,c(10,2:9)])
```

```{r}
rownames(pl2) <- c()
```

```{r}
pl2[1:5,]
```

```{r}
prem[1:5,]
```

```{r}
NationalResults <- function(nation) {

    data <- LoadAndScaleData(nation)
    results <- LeagueResults(data)
    return(results)
}

@@

```{r}
correct_res <- read.csv('../data/league_results/England17-18.csv')
```

```{r}
correct_res
```

```{r}
france <- NationalResults('France')
germany <- NationalResults('Germany')
italy <- NationalResults('Italy')
spain <- NationalResults('Spain')
```

```{r}
premiership <- NationalResults('England')
```

```{r}
ShowMatchEvents <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize events on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$location_y <- 68 - game$location_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to main events
    game_events <- game[game$eventName == 'Pass'| 
                        game$eventName == 'Shot' | 
                        game$eventName == 'On the Ball' | 
                        game$eventName == 'Challenge' | 
                        game$eventName == 'Free Kick',]

    # draw pitch
    p <- DrawPitch(theme='grey', arrow=direction_of_play) + 
        geom_point(data = game_events, 
               aes(location_x , location_y, fill=eventName, shape=eventName), pch=21, alpha=0.6, size=2 ) + 
        geom_point(data=game_events[game_events$subEventName=='Goal',], 
                   aes(location_x, location_y), shape=13, size=5) +
        geom_label_repel(data=game_events[game_events$subEventName=='Goal',], 
                         aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) +
        theme(legend.direction='horizontal', legend.position=c(0.5,0)) +
        scale_fill_manual(values=c("red", "black", 'white', "yellow", 'blue'), name='Event Type') +
        ggtitle(paste('Event Map:', team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

```{r}
#away vs Crystal Palace
match_id <- '2500032'
```

```{r}
# vs man city
match_id <- '2499943'
```

```{r}
ShowMatchEvents(events, match_id, home_or_away='Home')
ShowMatchEvents(events, match_id, home_or_away='Away', flip=T)
```

```{r}
ShowPassesAndShots <- function(events, match_id, team_name, home_or_away='Home', flip=F) {
    #' Visualize passes on pitch for given match and team.

    # select match
    game <- events[events$matchId==match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    # flip coordinates if desired
    if (flip==T) {
        game$location_x <- 105 - game$location_x
        game$target_x <- 105 - game$target_x
        game$location_y <- 68 - game$location_y
        game$target_y <- 68 - game$target_y
        direction_of_play = 'l'
    } else { 
        direction_of_play = 'r'
    }

    # limit attention to passes and shots
    passes <- game[game$eventName == 'Pass' & game$team == team_name,]
    shots <- game[(game$eventName=='Shot' | game$subEventName=='Goal') & game$team==team_name,]

    # draw pitch
    p <- DrawPitch(theme='grass', arrow=direction_of_play) + 
        geom_segment(data=na.exclude(passes), 
                     aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                     alpha=0.6, arrow = arrow(length = unit(0.1,"cm"))) +
        geom_label_repel(data=game[game$subEventName=='Goal',], 
                     aes(location_x, location_y, label = paste0(source,'(',matchPeriod,' ',time,')')), label.padding=0.1, size=2.3, alpha=1) + 
        geom_segment(data=na.exclude(shots), 
                 aes(x=location_x, y=location_y, xend=target_x, yend=target_y, color=subEventName), 
                 alpha=0.7, arrow = arrow(length = unit(0.1,"cm"))) +
        theme(legend.position=c(0.5,-0.01), legend.direction='horizontal',
              legend.background=element_rect(fill='#008000', linetype='solid')) +
        scale_color_manual(values=c("red", "black", 'blue', 'orange', "yellow", 'grey', 'purple', 'white','brown'), name='Pass Type') +
        ggtitle(paste('Pass Map:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')'))

    return(p)
} 
```

```{r}
ShowPassesAndShots(events, match_id, home_or_away='Home')
ShowPassesAndShots(events, match_id, home_or_away='Away', flip=T)
```

# Methodology

```{r}
PassNetwork <- function(events, match_id, team_name, team_colour='red', home_or_away='Home', flip=F, lower_threshold=1, high_threshold=10) {
    #' Draw Pass Map of First XI with nodes placed on mean (x,y) pitch-coordinates.

    game <- events[events$matchId == match_id,]
    goals <- game[game$subEventName=='Goal',]

    # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        game <- game[game$team==team_name,]
        home_or_away <- unique(game$home_or_away)
    } else {
        game <- game[game$home_or_away==home_or_away,]
        team_name <- unique(game$team)
    }

    # note opposition
    opposition <- teams[teams!=team_name]

    # get score
    team_score <- table(goals$team)[team_name]
    opposition_score <- table(goals$team)[opposition]

    firstXI <- game[game$FirstXI == 'True',]
    mean_positions <- firstXI[firstXI$location_x>0 & firstXI$location_y>0 & firstXI$location_x<105 & firstXI$location_y<68,] %>% 
        group_by(team, matchId, source) %>%
          dplyr::summarise(x_mean = mean(location_x), y_mean = mean(location_y)) %>% 
          ungroup() %>%
          mutate(team = as.factor(team), id = as.factor(matchId)) %>%
          as.data.frame()

    pass_counts <- ddply(data.frame(game$source, game$target),.(game.source,game.target),nrow)
    names(pass_counts) <- c('source','target','passcount')
    step1 <- merge(mean_positions, pass_counts, by='source')
    step2 <- step1[,c(1,4,5,7,8)]
    names(step2)[2:3] <- c('source_x','source_y')
    names(mean_positions)[3] <- 'target'
    step3 <- merge(mean_positions, step2, by='target')
    team <- step3[step3$team==team_name,]

    if (flip==T) {
        team$source_x <- 105 - team$source_x
        team$x_mean <- 105 - team$x_mean
        team$source_y <- 68 - team$source_y
        team$y_mean <- 68 - team$y_mean
        mean_positions$x_mean <- 105 - mean_positions$x_mean
        mean_positions$y_mean <- 68 - mean_positions$y_mean
    }

    return (DrawPitch() +
        geom_segment(data=team[team$passcount>=lower_threshold,], size=1, colour=team_colour,
                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_segment(data=team[team$passcount>=high_threshold,], size=1.5, colour='black',
                     aes(x=source_x, y=source_y, xend=x_mean, yend=y_mean, alpha=passcount)) +
        geom_label_repel(data = mean_positions[mean_positions$team==team_name,], 
                         aes(x_mean, y_mean, label = target), label.padding=0.1, size=2.3, alpha=1) +
        geom_point(data=team, aes(x_mean, y_mean,), fill=team_colour, colour='black', pch=21, size=3) +
        ggtitle(paste('Passing Network:',team_name,'playing',home_or_away,'versus',opposition,'(',team_score,'-',opposition_score,')')) +
        theme(legend.position=c(0.5,0.07), legend.direction='horizontal'))

```{r}
PassNetwork(events, match_id, team_colour='red')
PassNetwork(events, match_id, team_colour='blue', home_or_away='Away', flip=T)
```

```{r}
2499943
```

```{r}
GameGraph <- function(events, match_id, team_name, home_or_away='Home') {
    #' Return Possession Graph for given Match and Team.

    game <- events[events$matchId == match_id,]
    game <- game[game$source!='' & game$target!='' & game$target!='nan',]

     # note teams
    teams <- unique(game$team)

    # limit events to team specified by name or home/away
    if(!missing(team_name)){
        team_game <- game[game$team==team_name,]
        home_or_away <- unique(team_game$home_or_away)
    } else {
        team_game <- game[game$home_or_away==home_or_away,]
        team_name <- as.character(unique(team_game$team))
    }

    opposition_game <- game[game$team != team_name,]
    opposition_team <- as.character(unique(opposition_game$team))

    team_nodes <- unique(c(as.character(team_game$source),as.character(team_game$target)))
    team_edges <- data.frame(team_game$source, team_game$target)
    team_graph <- graph_from_data_frame(d=team_edges, vertices=team_nodes, directed=TRUE)
    team_adj <- as.matrix(team_graph[])

    possession_seqs <- unique(team_game$possession)

    seq_start <- c()

    for (pseq in possession_seqs){
        if (sum(team_game$possession==pseq)>1){
            player <- as.character(team_game[team_game$possession==pseq,][1,]$source)
            if (player!='' & player!='nan'){
                seq_start <- append(seq_start, player)
            }
        }
    }

    start_counts <- table(as.factor(seq_start))

    for (i in (2:(length(start_counts)-1))){
        team_adj[nrow(team_adj)-1,][names(start_counts[i])] <- as.numeric(start_counts[i])
    }

    opposition_status <- as.character(unique(opposition_game$home_or_away))
    opp_to_opp <- as.numeric(summary(game[game$team != team_name,]$to_team)[opposition_status])

    opp_nodes <- unique(c(as.character(opposition_game$source),as.character(opposition_game$target)))
    opp_edges <- data.frame(opposition_game$source, opposition_game$target)
    opp_graph <- graph_from_data_frame(d=opp_edges, vertices=opp_nodes, directed=TRUE)
    opp_adj <- as.matrix(opp_graph[])

    opp_to_opp <- sum(opp_adj[,ncol(opp_adj)-1])
    team_adj[(nrow(team_adj)-1),(ncol(team_adj)-1)] <- opp_to_opp

    team_status <- unique(team_game$home_or_away)
    opposition_own_goals <- as.numeric(nrow(game[game$home_or_away==opposition_status & game$to_team==paste(team_status,'Goal'),]))
    team_adj[(nrow(team_adj)-1),(ncol(team_adj))] <- opposition_own_goals

    team_graph <- graph_from_adjacency_matrix(team_adj, mode='directed')

    team_graph$team <- team_name
    team_graph$opposition <- opposition_team
    team_graph$status <- home_or_away

    return(team_graph)
}
```

```{r}
AdjacencyMatrix <- function(graph){
    #' Return Adjacency Matrix for Graph

    return(as.matrix(graph[]))
}
```

```{r}
TransitionMatrix <- function(graph){
    #' Return Transition Matrix for Graph considered as time-homogeneous Markov Process

    matrix <- as.matrix(graph[])

    for (i in 1:nrow(matrix)){
        matrix[i,] <- matrix[i,]/sum(matrix[i,])
    }

    matrix[nrow(matrix),ncol(matrix)] <- 1
    matrix[nrow(matrix),] <- rep(0, ncol(matrix))

    # treat 'Goal' as absorbtion state
    matrix[nrow(matrix),ncol(matrix)] <- 1

    return(matrix)
}
```

```{r}
ExpectedScoringTime <- function(transition_matrix){
    #' Calculate expected scoring time based on transition matrix.

    max_steps <- 10000
#     transition_matrix <- TransitionMatrix(GameGraph(events, match_id))
    hitting_times <- rep(0,(ncol(transition_matrix)-2))

    for (i in (1:(ncol(transition_matrix)-2))){

        state_probabilities <- matrix(NA,nrow=max_steps+1,ncol=ncol(transition_matrix),
                                      dimnames=list(0:max_steps,(ncol(transition_matrix)-1):0))
        vector <- rep(0,ncol(transition_matrix))

        vector[i] <- 1
        state_probabilities[1,] <- vector

        for ( kk in 1:max_steps ) {
            state_probabilities[kk+1,] <- t(transition_matrix)%*%state_probabilities[kk,]
        }

        probs <- diff(state_probabilities[,ncol(transition_matrix)])
        hitting_time <- sum(probs*seq_along(probs))

        hitting_times[i] <- hitting_time
    }

    names(hitting_times) <- names(transition_matrix[1,])[1:(nrow(transition_matrix)-2)]

    return(hitting_times)

```{r}
SeasonAdjacencyMatrix <- function(events, team) {

    players <- unique(events[events$team==team & events$source!='' & events$target!='',]$source)
    n <- length(players) + 2
    squad_adj <- matrix(0,nrow=n,ncol=n)
    rownames(squad_adj) <- c(as.character(players), 'Opposition', 'Goal')
    colnames(squad_adj) <- rownames(squad_adj)

    games <- unique(events[events$team==team,]$matchId)

    for (game in games){

        # get graph and adjacency matrix for game
        game_graph <- GameGraph(events, game, team)
        game_adj <- AdjacencyMatrix(game_graph)

        # add values of game adjacency matrix to full-season matrix
        for (source in rownames(game_adj)){
            for (target in colnames(game_adj)){
                squad_adj[source,target] <- squad_adj[source,target] + game_adj[source,target]
            }
        }
    }

    return(squad_adj)
}
```

```{r}
MarkovExpectedGoals <- function(events, season_adjacency_matrix) {

    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
    season_transition <- TransitionMatrix(season_graph)
    seasonXST <- ExpectedScoringTime(season_transition)
    seasonMXG <- 1/seasonXST
    normalize <- seasonMXG/(mean(seasonMXG))
    avg_events_per_game <- nrow(events)/380
    scale <- normalize * avg_events_per_game - avg_events_per_game
    sorted <- sort(scale, decreasing=T)

    return(sorted)
}
```

```{r}
GoalProbabilities <- function(season_adjacency_matrix) {

    season_graph <- graph_from_adjacency_matrix(season_adjacency_matrix)
    season_transition <- TransitionMatrix(season_graph)
    goal_probs <- sort(season_transition[,'Goal'],T)[2:nrow(season_transition)] * 100

    return(goal_probs)
}
```

```{r}
leagueMXG <- function(events){

    all_mXG <- c()
    teams <- unique(events$team)

    for (team in teams){

        season_adjacency_matrix <- SeasonAdjacencyMatrix(events, team)
        mXG <- MarkovExpectedGoals(events, season_adjacency_matrix)

        all_mXG <- append(all_mXG, mXG)

    }

    return(all_mXG)
}
```

```{r}
mXG <- c()
leagues = c('England','France','Germany','Italy','Spain')
for (league in leagues){
    events <- LoadAndScaleData(league)
    mXG <- append(mXG, leagueMXG(events))
}
```

```{r}
sort(mXG,T)
```

```{r}
g <- GameGraph(events, match_id, home_or_away='Home')
```

```{r}
MatrixHeatMap <- function(matrix, number_format='%.d', title='Matrix Heatmap'){
    #' Show Matrix as HeatMap

    hm <- pheatmap(matrix,color=brewer.pal(15,'YlGnBu'),cluster_rows=F,cluster_cols=F,legend=F,
                    display_numbers=T,number_format=number_format,fontsize_number=9,angle_col='315',
                    main=title)

    return(hm)
}
```

```{r}
MatrixHeatMap(
    AdjacencyMatrix(g)[1:11,1:11]

)
```

```{r}
MatrixHeatMap(
    AdjacencyMatrix(
        GameGraph(events, match_id, home_or_away='Away'))[1:11,1:11]
)
```

```{r}
MatrixHeatMap(
    TransitionMatrix(
        GameGraph(events, match_id, home_or_away='Home')), 
    number_format='%.2f',
    title=''
)
```

```{r}
MatrixHeatMap(
    TransitionMatrix(
        GameGraph(events, match_id, home_or_away='Away')), 
    number_format='%.2f',
    title=''
)
```

```{r}
GoalSeqs <- function(events, match_id) {
    match <- events[events$matchId==match_id,]
    goals <- match[match$subEventName=='Goal',]
    goal_seqs <- unique(goals$possession)   
    return(goal_seqs)
}
```

```{r}
TabulateSequence <- function(events, possession_sequence) {
    # Return a markdown table starting from the row before a possession sequence begins, and continuing until it ends.

    start <- min(events[events$possession == possession_sequence,]$X)
    stop <- max(events[events$possession == possession_sequence,]$X) + 1
    table <- events[start:stop,]
    df <- data.frame(table$matchPeriod, table$time, table$team, table$source, table$subEventName)
    names(df) <- c('Half', 'Time', 'Team', 'Event Description', 'Player')

    return(kable(df))
}
```

```{r}
sequence <- '2500032-91-Liv-0'
```

```{r}
TabulateSequence(events, sequence)
```

```{r}
SequenceOnPitch <- function(events, possession_sequence){
    #' Draw Possession Sequence on Pitch

    data <- events[events$possession == possession_sequence,]

    p <- DrawPitch(theme='grass') + 
        geom_label_repel(data = data, aes(location_x, location_y, label = paste(source,time)), label.padding=0.1, size=2.3, alpha=1) +
        geom_curve(data = data, aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), 
                   show.legend=FALSE, size=1, alpha = 1) +
        geom_point(data = data, aes(location_x , location_y, shape=eventName), size=2) + 
        ggtitle(paste0(nrow(data),'-part ', unique(data$team)[1], ' Possession Sequence ending with ', 
                       data[nrow(data),]$subEventName, ' by ', data[nrow(data),]$source))  +
        theme(legend.position=c(0.5,0.08), legend.direction='horizontal', legend.title=element_blank())

    return(p)
}
```

```{r}
SequenceOnPitch(events, sequence)
```

youtube clip
<img src="../figures/youtube_clip.gif" width='700'>

```{r}
SequenceGraph <- function(events, possession_sequence){
    #' Return Graph of Possession Sequence

    data <- events[events$possession == possession_sequence,]
    data <- data[data$source!='' & data$target!='nan',]
    nodes <- unique(c(as.character(data$source),as.character(data$target)))
    edges <- data.frame(data$source, data$target)
    g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
    g$id <- possession_sequence
    g$data <- data
    g$team <- as.character(unique(data$team))

    return(g)
}
```

```{r}
VisualizeGraph <- function(possession_sequence_graph) {
    #' Visualize Possession Sequence as Linear Graph with Looping Edges

    g <- possession_sequence_graph
    data <- possession_sequence_graph$data

    visualization <- ggraph(g, 'linear') + 
        geom_edge_arc(aes(color=data$eventName), 
                      arrow=arrow(length=unit(4,'mm')), 
                      fold=F,
                     width=1) +
        geom_edge_loop(aes(color=data$eventName),
                      width=1) +
        geom_node_point(color='black', 
                        size=2,
                        alpha=0.5) +
        geom_node_text(aes(label = name), 
                        repel=T, 
                        angle=90, hjust=2, ) +
        scale_edge_colour_manual(
            values=c('indianred3', 'wheat4', 'grey', 'grey30',
                    'red','blue','green','orange','purple','brown','pink'
                    ),
            name='Event Type') +
        theme_void()

    return(visualization)

```{r}
g <- SequenceGraph(events, sequence)
```

```{r}
VisualizeGraph(g)
```

```{r}
GetGraphMetrics <- function(possession_graph) {
    #' Return basic metrics for given Graph

    g <- possession_graph
    path_metrics <- c(g$id, g$team, length(V(g)), length(E(g)), 
                      diameter(g), mean_distance(g), edge_density(g, loops=T), 
                      girth(g)$girth, transitivity(g), authority_score(g)$value, hub_score(g)$value)
    names(path_metrics) <- c('Label', 'Team', 'Nodes', 'Edges', 'Diameter', 'Mean_Distance', 
                             'Edge_Density', 'Girth', 'Transitivity', 'Authority_Score', 'Hub_Score')

    return(path_metrics)
}
```

```{r}
GetNodeMetrics <- function(possession_graph) {
    #' Return metrics for each Node of given Graph

    g <- possession_graph
    df <- data.frame(page_rank(g)$vector)
    names(df) <- 'pagerank'
    df$in_degree <- degree(g, mode='in')
    df$out_degree <- degree(g, mode='out')
    df$eccentricity <- eccentricity(g)
    df$closeness <- closeness(g)
    df$betweenness <- betweenness(g)
    df$norm_betweenness <- betweenness(g, normalized=T)
    df$alpha_centrality <- alpha_centrality(g)
    df$power_centrality <- power_centrality(g)
    df$eigen_centrality <- eigen_centrality(g)$vector
    df$authority <- authority_score(g)$vector
    df$hub <- hub_score(g)$vector

    return(df)
}
```

```{r}
GetEdgeMetrics <- function(possession_graph) {
    #' Return metrics for each Edge of given Graph

    g <- possession_graph
    df <- data.frame(as_edgelist(g))
    names(df) <- c('Source','Target')
    df$sequence <- 1:length(E(g))
    df$edge_betweenness <- edge_betweenness(g)

    return(df)
}
```

```{r}
GetGraphMetrics(g)
```

```{r}
GetNodeMetrics(match_graph)
```

```{r}
GetEdgeMetrics(g)
```

```{r}
sequence2 <- '2499858-18-Liv-5'
```

```{r}
TabulateSequence(sequence2)
```

```{r}
SequenceOnPitch(sequence2)
```

<img src='../figures/mane_vs_stoke.gif'>

```{r}
g2 <- SequenceGraph(sequence2)
VisualizeGraph(g2)
```

```{r}
GetGraphMetrics(g2)
```

```{r}
sequence3 <-  '2499889-18-Liv-0'
```

```{r}
SequenceOnPitch(sequence3)
```

```{r}
g3 <- SequenceGraph(sequence3)
VisualizeGraph(g3)
```

```{r}
GetGraphMetrics(g3)
```

<img src="../figures/youtube_clip2.gif">

>>>>

---
jupyter:
  jupytext:
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.4.2
  kernelspec:
    display_name: R
    language: R
    name: ir
---

```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(MASS)
library(cowplot)
library(forcats)
library(ggforce)
library(ggrepel)
library(scales)
library(tidyr)
library(xts)
library(zoo)
```

```{r}
library(ggraph)
```

```{r}
library(igraph)
```

```{r}
soccerPitch <- function(lengthPitch = 105, widthPitch = 68, arrow = c("none", "r", "l"), title = NULL, subtitle = NULL, theme = c("light", "dark", "grey", "grass")) {

@@

  # outer border (t,r,b,l)
#   border <- c(10, 6, 5, 6)
  border <- c(5, 5, 5, 5)

@@

#   # add arrow
#   if(arrow[1] == "r") {
#     p <- p + 
#       geom_segment(aes(x = 0, y = -2, xend = lengthPitch / 3, yend = -2), colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
#   } else if(arrow[1] == "l") {
#     p <- p + 
#       geom_segment(aes(x = lengthPitch, y = -2, xend = lengthPitch / 3 * 2, yend = -2), colour = arrowCol, size = 1.5, arrow = arrow(length = unit(0.2, "cm"), type="closed"), linejoin='mitre')
#   }

#   # add title and/or subtitle
#   theme_buffer <- ifelse(theme[1] == "light", 0, 4)
#   if(!is.null(title) & !is.null(subtitle)) {
#     p <- p +
#       cowplot::draw_text(title, 
#                          x = 0, y = widthPitch + 9, hjust = 0, vjust = 1,
#                          size = 15, fontface = 'bold', col = colText) +
#       cowplot::draw_text(subtitle, 
#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
#                          size = 13, col = colText) +
#       theme(plot.margin = unit(c(-0.525,-0.9,-0.7,-0.9), "cm"))
#   } else if(!is.null(title) & is.null(subtitle)) {
#     p <- p +
#       cowplot::draw_text(title, 
#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
#                          size = 15, fontface = 'bold', col = colText) +
#       theme(plot.margin = unit(c(-0.9,-0.9,-0.7,-0.9), "cm"))
#   } else if(is.null(title) & !is.null(subtitle)) {
#     p <- p +
#       cowplot::draw_text(subtitle, 
#                          x = 0, y = widthPitch + 4.5, hjust = 0, vjust = 1,
#                          size = 13, col = colText) +
#       theme(plot.margin = unit(c(-0.9,-0.9,-0.7,-0.9), "cm"))
#   } else if(is.null(title) & is.null(subtitle)){
#     p <- p +
#       theme(plot.margin = unit(c(-1.2,-0.9,-0.7,-0.9), "cm"))
#   }

```{r}
events <- read.csv('../data/processed_events_England.csv')

@@

events$location_y <- y_scale - (events$location_y/100 * y_scale)
events$target_y <- y_scale - (events$target_y/100 * y_scale)
```

```{r}
game <- events[events$matchId == '2500032',]
```

```{r}
game_events <- game[game$eventName == 'Pass'| game$eventName == 'Shot' | game$eventName == 'Duel' | game$eventName == 'Free Kick',]
```

```{r}
soccerPitch(theme='grey') + 
    geom_point(data = game_events, 
           aes(location_x , location_y, fill=eventName), pch=21, alpha=0.5, size=2 ) + 
    theme(legend.position='bottom', legend.title = element_blank()) +
    scale_fill_manual(values=c("red", "black", "blue", 'white'))
```

```{r}
id <-  '2500032-91-Liv-0'
# id <- best_goal_sequences[1]
data <- events[events$possession == id,]
# 534 is brighton vs bournemouth
# data <- events[events$possession == '2499759-46-Bri&HovAlb-0',]
# data <- events[events$possession == '2499889-18-Liv-0',]
# data <- events[events$possession == '2499774-85-ManCit-0',] #2500065-23-ManCit-1',]#2499913-77-Liv-0',]

soccerPitch(theme='grass') + geom_point(data = data,
               aes(location_x , location_y)) + 
    geom_label_repel(data = data, aes(location_x, location_y, label = source), label.padding=0.1, size=2.3, alpha=1) +
 geom_curve(data = data,
              aes(x = location_x, xend = target_x, y = location_y, yend = target_y*.99, col = X), show.legend=FALSE, size=1, alpha = 1)

@@

<img src="../figures/youtube_clip.gif">

```{r}
nodes <- unique(c(as.character(data$source),as.character(data$target)))
edges <- data.frame(data$source, data$target)
g <- graph_from_data_frame(d=edges, vertices=nodes, directed=TRUE)
ggraph(g, layout='linear') + 
    geom_edge_arc(aes(start_cap = label_rect(node1.name), 
                       end_cap = label_rect(node2.name)), 
                  arrow=arrow(length=unit(4,'mm')), 
                  color='grey', width=0.7, fold=F) +
    geom_edge_loop(color='grey', linetype='dashed') +
    geom_node_point(size=2, color='black', fill='white') +
    geom_node_text(aes(label = name), repel=T, angle=90, hjust=2, ) +
    theme_void()
```

@@
